---
title: "Indexation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Indexation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo=FALSE, message=FALSE}
library(ilabelled)
```


# Indexation

One of the fundamental features of i_labelled objects is that both underlying values and value labels can be used for logical operations. 

The following rules must be observed when indexing i_labelled objects: 

  1. If numerical values are used for the indexation of numerical i_labelled data, the underlying numerical values of the i_labelled data are used for the comparison. 
  2. If character values or values of class i_labelled are used for indexation, *value labels are always used over the underlying values of the i_labelled data*.


## Example 1: Indexation of unterlying data

```{r}
myData <- i_labelled(
  x = c(1, 2, 3, NA), 
  labels = c("A" = 1, "B" = 2, "C" = 3), 
  label = "my Variable"
)

myData %in% 1
myData == 1
myData > 1
```


## Example 2: Indexation of value labels

Keep in mind, if character values are used for indexation, value labels are alway chosen over underlying values.

```{r}
myData %in% c("A", "B")
myData != "C"
myData > "A"
```


